---
title: "INET4061-Lab2"
author: "Mahsa Ayoughi"
date: "9/19/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
Goals: Create a R markdown and plot the statistical functions, PDF, CDP, Inverse CDF, Normal Distribution and t distribution.
```{r imports}
require(ggplot2)
require(grid)
require (RCurl)
require(graphics)
```


## Data 
Create and plot the following functions: Probability Density Function, Cumulative Distribution Function (CDF), Quantile or Inverse CDF, Normal Distribution and t Distribution.

### Probability Density Function 
The PDF is the density of probability rather than the probability mass. The concept is very similar to mass density in physics: its unit is probability per unit length.


```{r}
mean = 0
sd=1
x <- seq(-4, 4, length=100)
PDF <- function (x, type= "norm", mean, sd){
    y <- dnorm(x, mean, sd)
    ggplot(x,y) 
    
}
```

### Cumulative Distribution Function

The CDF determine the probability that an observation drawn from a distribution falls below a specified value. In essence, cdf() gives you the area under the distribution's density curve to the left of a certain value on the x axis.
```{r}
x <- seq(-4, 4, length=100)
CDF <- function (x, mean, sd){
  y <- pnorm(x, mean, sd)
  plot2 <- ggplot (x,y)
  print (plot2)
}
```


### Inverse Cumulative Density Function
This gives the inverse of the CDF where it points to the opposite.
This graph increases quickly, then plateaus for the rest. 

```{r}
ICDF <- function (x, mean, sd){
  y <- qnorm(x, mean, sd)
  plot3 <- ggplot (x,y)
  print (plot3)
}
```

### T Distribution
Code Resource: https://stat.ethz.ch/R-manual/R-devel/library/stats/html/TDist.html
```{r}

1 - pt(1:5, df = 1)
qt(.975, df = c(1:4,8,12,24))
tt <- seq(0, 10, len = 21)
ncp <- seq(0, 6, len = 31)
ptn <- outer(tt, ncp, function(t, d) pt(t, df = 3, ncp = d))
t.tit <- "Non-central t - Probabilities"
image(tt, ncp, ptn, zlim = c(0,1), main = t.tit)


```



## Conclusions 
Question 2: What is the probability that the Z-score is within two standard deviations of the mean?
Approximately 95% of the z-score will be within two standard deviations of the mean.

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
